Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-06-10 thirdPaper/draft/typeConeAssociahedra.tex
changes Parts 1 and 2

% geometry
\newcommandx{\Asso}[2][1=n,2={}]{\mathsf{Asso}^{#2}(#1)} % associahedron
\newcommandx{\Zono}[2][1=n,2={}]{\mathsf{Zono}^{#2}(#1)} % zonotope
\newcommand{\Fan}{\mathcal{F}} % fan
\newcommand{\multiplicityVector}{\b{m}} % multiplicity vector
\newcommandx{\ray}[1][1=r]{\b{#1}} % ray
\newcommandx{\rays}[1][1=R]{\b{#1}} % rays
\newcommand{\gvector}[1]{\b{g}(#1)} % g-vector of #1
\newcommand{\gvectorFull}[2]{\b{g}(#1,#2)} % g-vector of #2 wrt #1
\newcommand{\gvectors}[1]{\b{g}(#1)} % g-vectors of #1
\newcommand{\gvectorsFull}[2]{\b{g}(#1,#2)} % g-vectors of #2 wrt #1
\newcommandx{\gvectorFan}[1][1=\quiver]{\mathcal{F}(#1)} % g-vector fan
\newcommand{\cvector}[2]{\mathbf{c}(#2 \in #1)} % c-vector of the cluster variable #2 in the cluster #1
\newcommand{\cvectorFull}[3]{\mathbf{c}(#1,#3 \in #2)} % c-vector of the cluster variable #3 in the cluster #2 with respect to the initial cluster #1
\newcommand{\cvectors}[1]{\mathbf{c}(#1)} % c-vectors of the cluster #1
\newcommand{\cvectorsFull}[2]{\mathbf{c}(#1,#2)} % c-vectors of the cluster #2 with respect to the initial cluster #1

@@

\vspace{2cm}
\centerline{\includegraphics[scale=.8]{typeConeCoarsenings}}
\vspace{.5cm}
\centerline{The type cone of the cluster fan of type~$A_2$.}

@@

\smallskip
\item Once we get a representation of a polytope~$P$ as a signed Minkowski sum~$P = \sum_{i \in [k]}\alpha_i Q_i$, we get a formula for the volume of $P$ as a multivariate polynomial in $\alpha_1,\dots,\alpha_k$ whose coefficients are the mixed volumes of $Q_1,\dots, Q_k$~\cite{ArdilaBenedettiDoker,McMullen-Valuations}. These encode rich combinatorial information when representing generalized permutahedra as positive Minkowski sums of coordinate simplices or hypersimplices~\cite{Postnikov}, and in particular for matroid polytopes~\cite{ArdilaBenedettiDoker}. It would be interesting to get such volume formulas for polytopal realizations of $\b{g}$-vector fans of cluster algebras and gentle algebras, and combinatorial interpretations for their coefficients.
\end{itemize}

\smallskip
\noindent
Some of these questions are explicitly indicated throughout the paper.
To conclude, we observe that the associahedra constructed in~\cite{ArkaniHamedBaiHeYan, BazierMatteDouvilleMousavandThomasYildirim} and the present paper describe all polytopal realizations of the $\b{g}$-vector fan.
This motivates the following question: Is it possible to similarly understand the complete realization space of the cluster complex, including all realizations not having the~$\b{g}$-vector fan as their normal fan?
\pg{AWESOME introduction.}
\vincent{:-)}

@@

\begin{notation}
For any two adjacent cones~$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan$ with~$\rays \ssm \{\ray\} = \rays' \ssm \{\ray'\}$, we denote by~$\coefficient[{\ray[s]}][\rays][\rays']$ the coefficient of~$\ray[s]$ in the unique linear dependence between the rays of~$\rays \cup \rays'$, \ie such that
\[
\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \ray[s] = 0.
\]
These coefficients are \apriori{} defined up to rescaling, but we additionally fix the rescaling so that~${\coefficient[\ray][\rays][\rays'] + \coefficient[\ray'][\rays][\rays'] = 2}$ (this convention is arbitrary, but will be convenient in \cref{sec:applications}).
\end{notation}

@@

\begin{definition}
\label{def:typeCone}
The \defn{type cone} of~$\Fan$ is the cone~$\typeCone(\Fan)$ of all $\Fan$-admissible height vectors~$\b{h}$:
\begin{align*}
\typeCone(\Fan) & \eqdef \set{\b{h} \in \R^N}{\Fan \text{ is the normal fan of } P_\b{h}} \\
& = \Bigset{\b{h} \in \R^N}{\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} > 0 \; \begin{array}{l} \text{for any adjacent maximal} \\ \text{cones~$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan$} \end{array}}.
\end{align*}
\end{definition}

@@

We denote by $\ctypeCone(\Fan)$ the closure of $\typeCone(\Fan)$, and call it the \defn{closed type cone} of~$\Fan$. It is the closed polyhedral cone defined by the inequalities $\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} \geq 0$ for adjacent maximal cones~$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$. If $\Fan$ is the normal fan of the polytope~$P$, then $\ctypeCone(\Fan)$ is called the \defn{deformation cone} of $P$ in~\cite{Postnikov}.

@@

\begin{definition}
An \defn{extremal adjacent pair} of~$\Fan$ is a pair of adjacent maximal cones~$\{\R_{\ge0}\rays, \R_{\ge0}\rays'\}$ of~$\Fan$ such that the corresponding inequality $\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} \geq 0$ in the definition of the type cone~$\ctypeCone(\Fan)$ actually defines a facet of~$\ctypeCone(\Fan)$.
\end{definition}

@@

\begin{remark}
\label{rem:dimTypeCone}
Since the type cone is an~$N$ dimensional cone with a linearity subspace of dimension~$n$, it has at least~$N-n$ facets.
We say that the type cone is \defn{simplicial} when it has precisely $N-n$ facets.
\end{remark}

@@

\begin{proposition}
\label{prop:non-simplicial}
Let $\c{G}$ be a complete fan that coarsens the essential complete simplicial fan~$\Fan$.
The \defn{type cone}~$\typeCone(\c{G})$ of all $\c{G}$-admissible height vectors~$\b{h}$ is
\begin{align*}
\typeCone(\c{G}) 
& = \left\{\b{h} \in \R^N\;\middle|\;
\begin{array}{ll}
\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} = 0
& \begin{array}{l} \text{for any adjacent maximal cones} \\ \text{$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan$ belonging}\\\text{to \textbf{the same} maximal cone of $\c{G}$,}\end{array}
\\[.7cm]
\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} > 0
& \begin{array}{l} \text{for any adjacent maximal cones} \\ \text{$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan$ belonging}\\\text{to \textbf{distinct} maximal cones of $\c{G}$}\end{array}
\end{array}
\right\}.
\end{align*}
\end{proposition}

@@

\begin{definition}
\label{def:uerp}
We say that two exchangeable rays~$\ray, \ray'$ of~$\Fan$ admit a \defn{unique exchange relation} when the linear dependence
\[
\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \ray[s] = 0.
\]
does not depend on the pair~$\{\rays, \rays'\}$ of adjacent maximal cones of~$\Fan$ with $\rays \ssm \{\ray\} = \rays' \ssm \{\ray'\}$, but only on the pair of rays~$\ray, \ray'$.
This implies in particular that the rays~$\ray[s]$ for which~$\coefficient[{\ray[s]}][\rays][\rays'] \ne 0$ belong to~$\rays \cup \rays'$ for any pair~$\{\rays, \rays'\}$ of adjacent cones of~$\Fan$ with $\rays \ssm \{\ray\} = \rays' \ssm \{\ray'\}$.

@@

\begin{definition}
In a fan~$\Fan$ with the unique exchange relation property, an \defn{extremal exchangeable pair} is a pair of exchangeable rays~$\{\ray,\ray'\}$ such that the corresponding inequality~${\sum_{\ray[s]} \coefficient \, \b{h}_{\ray[s]} \geq  0}$ defines a facet of the closed type cone~$\ctypeCone(\Fan)$.
\end{definition}

@@

We still consider an essential complete simplicial fan~$\Fan$ in~$\R^n$, the $N \times n$-matrix~$\b{G}$ whose rows are the rays of~$\Fan$, and the $(N-n) \times N$-matrix~$\b{K}$ which spans the left kernel of~$\b{G}$.

@@

\begin{corollary}
\label{coro:simplicialTypeCone}
Assume that the type cone~$\typeCone(\Fan)$ is simplicial and let~$\b{K}$ be the $(N-n) \times N$-matrix whose rows are the outer normal vectors of the facets of~$\typeCone(\Fan)$. Then the polytope
\[
R_\b{\ell} \eqdef \set{\b{z} \in \R^N}{K\b{z} = \b{\ell} \text{ and } \b{z} \ge 0}
\]
is a realization of the fan~$\Fan$ for any positive vector~$\b{\ell} \in \R_{>0}^{N-n}$.
Moreover, the polytopes~$R_\b{\ell}$ for~$\b{\ell} \in \R_{>0}^{N-n}$ describe all polytopal realizations of~$\Fan$.
\end{corollary}

\begin{proof}
Let~$\b{\ell} \in \R_{>0}^{N-n}$.
Since~$\b{K}$ has full rank there exists~$\b{h} \in \R^N$ such that~$\b{K}\b{h} = \b{\ell}$.
Since~$\b{K}\b{h} \ge 0$ and the rows of~$\b{K}$ are precisely all outer normal vectors of the facets of the type cone~$\typeCone(\Fan)$, we obtain that~$\b{h}$ belongs to~$\typeCone(\Fan)$.
Since $R_\b{\ell} = Q_\b{h} \sim P_\b{h}$ by \cref{prop:alternativePolytopalRealization}, we conclude that~$R_\b{\ell}$ is a polytopal realization of~$\Fan$.
Since~$\typeCone(\Fan)$ is simplicial, we have~$\b{K}\typeCone(\Fan) = \R_{>0}^{N-n}$, so that we obtain all polytopal realizations of~$\Fan$ this way.
\end{proof}

@@

We say that~$P$ is a \defn{Minkowski summand} of~$R$ if there is a polytope~$Q$ such that $R=P+Q$, and a \defn{weak Minkowski summand} if there is a $\lambda \ge 0$ and a polytope~$Q$ such that ${\lambda R = P + Q}$. 
\cref{lem:MinkowskiSum} ensures that convex combinations in the type cone correspond to Minkowski combinations of polytopes. Actually, the set of weak Minkowski summands of a polytope $P$ with normal fan~$\Fan$ is isomorphic to the closed type cone~$\ctypeCone(\Fan)$ modulo linearity (the cone of weak Minkowski summands was studied by W.~Meyer in~\cite{Meyer}, to see that it is equivalent to the type cone and other formulations, see for example the appendix of~\cite{PostnikovReinerWilliams}).

@@

If $P = Q + R$, then we say that~$R$ is the \defn{Minkowski difference} of~$P$ and~$Q$, denoted~${R = P - Q}$. Note that this is only defined when~$Q$ is a Minkowski summand of~$P$.
There are two natural ways this construction generalizes to arbitrary pairs of polyhedra. On the one hand, we have $P - Q = \set{x \in \R^n}{x + Q \subseteq P}$. 
Notice that $P_\b{h} - P_{\b{h}'} = P_{\b{h} - \b{h'}}$ when $P_{\b{h}'}$ is a Minkowski summand of $P_\b{h}$. Considering arbitrary differences of support functions gives rise to \defn{virtual polyhedra}, which is the Grothendieck group of the semigroup of polytopes with Minkowski sums~\cite{PukhlikovKhovanskii}.

@@

\begin{corollary}
If the type cone~$\typeCone(\Fan)$ of $\Fan$ is simplicial, then the $N-n$ idecomposable polytopes $P_{\b{h}^{(1)}}, \dots, P_{\b{h}^{(N-n)}}$ arising from the rays of $\ctypeCone(\Fan)$ verify that 
\begin{enumerate}[(i)]
\item  every $P_\b{h}$ with $\b{h}\in \typeCone(\Fan)$ has a unique representation (up to translation) as a Minkowski sum of positive dilates of $P_{\b{h}^{(1)}}, \dots, P_{\b{h}^{(N-n)}}$; and
\item  every $P_\b{h}$ with $\b{h}\in \ctypeCone(\Fan)$ has a unique representation (up to translation) as a Minkowski sum of non-negative dilates of $P_{\b{h}^{(1)}}, \dots, P_{\b{h}^{(N-n)}}$.
\end{enumerate} 
\end{corollary}

@@

\begin{example}
\label{exm:typeConeMinkowskiSum}
\enlargethispage{.4cm}
Consider the $2$-dimensional fan~$\Fan$ of \cref{fig:typeCone}\,(left) whose type cone has been described in \cref{exm:typeCone} and represented in \cref{fig:typeCone}\,(right).
As illustrated in \cref{fig:typeConeCoarsenings}, the rays of the closed type cone~$\ctypeCone(\Fan)$ are directed by the height vectors~$(1,2,1,2,1)$ (corresponding to a triangle), $(0,1,1,0,1)$ (corresponding to a vertical segment) and~$(1,0,1,1,0)$ (corresponding to a horizontal segment).
Therefore, any polytope in~$\ctypeCone(\Fan)$ is (up to translation) a positive Minkowski combination of a triangle and two edges.
\cref{fig:typeConeMinkowskiSum} illustrates this property for the polytope of \cref{fig:typeCone}\,(middle) corresponding to the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$.
Note that~$\Fan$ is the normal fan of J.-L.~Loday's $2$-dimensional associahedron, whose realization as a Minkowski sum of faces of a simplex was given by A.~Postnikov in~\cite{Postnikov}.

@@

\begin{theorem}
\label{thm:associahedronLoday}
The associahedron~$\Asso[n]$ can be described equivalently as:
\begin{itemize}
\item the convex hull of the points~$L(T) \in \R^{n+1}$ for all rooted binary trees~$T$ with~$n+1$ internal nodes, where the $i$th coordinate of~$L(T)$ is the product of the number of leaves in the left subtree by the number of leaves in the right subtree of the $i$th node of~$T$ in inorder~\cite{Loday},
\item the intersection of the hyperplane~$\bigset{\b{x} \in \R^{n+1}}{\sum_{\ell \in [n+1]} x_\ell = \binom{n+2}{2}}$ with the halfspaces $\bigset{\b{x} \in \R^{n+1}}{\sum_{a < \ell < b} x_\ell \ge \binom{b-a+2}{2}}$ for all internal diagonal~$(a,b)$ of \mbox{the $(n+3)$-gon~\cite{ShniderSternberg}},
\item the Minkowski sum of the faces~$\simplex_{[i,j]}$ of the standard $n$-dimensional simplex~$\simplex$ corresponding to all proper intervals~$[i,j]$ of~$[n+1]$,~\cite{Postnikov}.
\end{itemize}
\end{theorem}

@@

\begin{corollary}
\label{coro:typeConeAsso}
Let~$n \in \N$ and~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$. Then the type cone of the normal fan~$\gvectorFan[n]$ of~$\Asso[n]$ is given by
\[
\typeCone(\gvectorFan[n]) = \set{\b{h} \in \R^{X(n)}}{\begin{array}{l} \b{h}_{(0,n+2)} = 0, \qquad\text{and}\qquad \b{h}_{(a,a+1)} = 0 \text{ for all } 0 \le a \le n+1 \\ \b{h}_{(a,b)} + \b{h}_{(a',b')} > \b{h}_{(a,b')} + \b{h}_{(a',b)} \text{ for all } 0 \le a < a' < b < b' \le n+2 \end{array}}.
\]
\end{corollary}

@@

\arnau{I think the coordinates of the g-vector corresponding to {diagonalA3} were wrong, and I swapped them. }

\arnau{When writing the coordinates of the g's, we cannot follow \cref{thm:normalFanLoday}, because we don't say which basis have we chosen for the hyperplane.}
\medskip
\noindent
In all our tables, we just record the coefficients of the inequalities, which are then easily reconstructed.
For instance, the inequality~$\red \circled{A}$ above is given by~$\b{h}_{\includegraphics[scale=.3]{diagonalA1}} + \b{h}_{\includegraphics[scale=.3]{diagonalA3}} > \b{h}_{\includegraphics[scale=.3]{diagonalA2}}$.

The type cone of~$\gvectorFan[2]$ lives in~$\R^9$ and has a linearity space of dimension~$3$.
It has $6$ facet-defining inequalities (given below), which correspond to the flips described in \cref{prop:extremalExchangeablePairsAsso} and illustrated in \cref{fig:labelFacetDefiningInequalititiesAsso}\,(left).

@@

\arnau{Change $\b{f}_{(a,b)}$ to $\b{h}_{(a,b)}$? (To homogenize with previous sections)}

@@

\begin{corollary}
\label{coro:simplicialTypeConeAsso}
The type cone~$\typeCone(\gvectorFan[n])$ is simplicial.
\end{corollary}

@@

\para{Mutations in the cluster fan}
%
We now discuss the linear dependences between $\b{g}$-vectors of adjacent seeds.
The following proposition was stated in~\cite[Lem.~19]{HohlwegPilaudStella}.

>>>>

% geometry
\newcommandx{\Asso}[2][1=n,2={}]{\mathsf{Asso}^{#2}(#1)} % associahedron
\newcommandx{\Zono}[2][1=n,2={}]{\mathsf{Zono}^{#2}(#1)} % zonotope
\newcommandx{\Fan}[1][1=F]{\mathcal{#1}} % fan
\newcommand{\multiplicityVector}{\b{m}} % multiplicity vector
\newcommandx{\ray}[1][1=r]{\b{#1}} % ray
\newcommandx{\rays}[1][1=R]{\b{#1}} % rays
\newcommand{\gvector}[1]{\b{g}(#1)} % g-vector of #1
\newcommand{\gvectorFull}[2]{\b{g}(#1,#2)} % g-vector of #2 wrt #1
\newcommand{\gvectors}[1]{\b{g}(#1)} % g-vectors of #1
\newcommand{\gvectorsFull}[2]{\b{g}(#1,#2)} % g-vectors of #2 wrt #1
\newcommandx{\gvectorFan}[1][1=\quiver]{\mathcal{F}(#1)} % g-vector fan
\newcommand{\cvector}[2]{\mathbf{c}(#2 \in #1)} % c-vector of the cluster variable #2 in the cluster #1
\newcommand{\cvectorFull}[3]{\mathbf{c}(#1,#3 \in #2)} % c-vector of the cluster variable #3 in the cluster #2 with respect to the initial cluster #1
\newcommand{\cvectors}[1]{\mathbf{c}(#1)} % c-vectors of the cluster #1
\newcommand{\cvectorsFull}[2]{\mathbf{c}(#1,#2)} % c-vectors of the cluster #2 with respect to the initial cluster #1

@@

\vspace{1.5cm}
\centerline{\includegraphics[scale=.8]{typeConeCoarsenings}}
\vspace{.5cm}
\centerline{The type cone of the cluster fan of type~$A_2$.}

@@

\smallskip
\item Once we get a representation of a polytope~$P$ as a signed Minkowski sum~$P = \sum_{i \in [k]} \alpha_i Q_i$, we get a formula for the volume of $P$ as a multivariate polynomial in $\alpha_1, \dots, \alpha_k$ whose coefficients are the mixed volumes of $Q_1, \dots, Q_k$~\cite{ArdilaBenedettiDoker, McMullen-Valuations}. These encode rich combinatorial information when representing generalized permutahedra as positive Minkowski sums of coordinate simplices or hypersimplices~\cite{Postnikov}, and in particular for matroid polytopes~\cite{ArdilaBenedettiDoker}. It would be interesting to get such volume formulas for polytopal realizations of $\b{g}$-vector fans of cluster algebras and gentle algebras, and combinatorial interpretations for their coefficients.
\end{itemize}

\smallskip
\noindent
Some of these questions are explicitly indicated throughout the paper.
To conclude, we observe that the associahedra constructed in~\cite{ArkaniHamedBaiHeYan, BazierMatteDouvilleMousavandThomasYildirim} and the present paper describe all polytopal realizations of the $\b{g}$-vector fan.
This motivates the following question: Is it possible to similarly understand the complete realization space of the cluster complex, including all realizations not having the~$\b{g}$-vector fan as their normal fan?

@@

\begin{notation}
For any adjacent maximal cones~$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan$ with ${\rays \ssm \{\ray\} = \rays' \ssm \{\ray'\}}$, we denote by~$\coefficient[{\ray[s]}][\rays][\rays']$ the coefficient of~$\ray[s]$ in the unique linear dependence between the rays of~$\rays \cup \rays'$, \ie such that
\[
\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \ray[s] = 0.
\]
These coefficients are \apriori{} defined up to rescaling, but we additionally fix the rescaling so that~${\coefficient[\ray][\rays][\rays'] + \coefficient[\ray'][\rays][\rays'] = 2}$ (this convention is arbitrary, but will be convenient in \cref{sec:applications}).
\end{notation}

@@

\begin{definition}
\label{def:typeCone}
The \defn{type cone} of~$\Fan$ is the cone~$\typeCone(\Fan)$ of all $\Fan$-admissible height vectors~$\b{h} \in \R^N$:
\begin{align*}
\typeCone(\Fan) & \eqdef \set{\b{h} \in \R^N}{\Fan \text{ is the normal fan of } P_\b{h}} \\
& = \Bigset{\b{h} \in \R^N}{\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} > 0 \; \begin{array}{l} \text{for any adjacent maximal} \\ \text{cones~$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan$} \end{array}}.
\end{align*}
\end{definition}

@@

We denote by $\ctypeCone(\Fan)$ the closure of $\typeCone(\Fan)$, and call it the \defn{closed type cone} of~$\Fan$. It is the closed polyhedral cone defined by the inequalities $\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} \ge 0$ for any adjacent maximal cones~$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$. If $\Fan$ is the normal fan of the polytope~$P$, then $\ctypeCone(\Fan)$ is called the \defn{deformation cone} of $P$ in~\cite{Postnikov}, see also~\cite{PostnikovReinerWilliams}.

@@

\begin{definition}
An \defn{extremal adjacent pair} of~$\Fan$ is a pair of adjacent maximal cones~$\{\R_{\ge0}\rays, \R_{\ge0}\rays'\}$ of~$\Fan$ such that the corresponding inequality $\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} \ge 0$ in the definition of the type cone~$\ctypeCone(\Fan)$ actually defines a facet of~$\ctypeCone(\Fan)$.
\end{definition}

@@

\begin{remark}
\label{rem:dimTypeCone}
Since the type cone is an~$N$ dimensional cone with a linearity subspace of dimension~$n$, it has at least~$N-n$ facets (thus~$N-n$ extremal adjacent pairs).
We say that the type cone is \defn{simplicial} when it has precisely $N-n$ facets.
\end{remark}

@@

\begin{proposition}
\label{prop:non-simplicial}
Let $\Fan$ be a complete fan that coarsens the essential complete simplicial fan~$\Fan[G]$.
The \defn{type cone}~$\typeCone(\Fan)$ of all $\Fan$-admissible height vectors~$\b{h} \in \R^N$ is
\begin{align*}
\typeCone(\Fan) 
& = \left\{\b{h} \in \R^N\;\middle|\;
\begin{array}{ll}
\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} = 0
& \begin{array}{l} \text{for any adjacent maximal cones} \\ \text{$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan[G]$ belonging}\\\text{to \textbf{the same} maximal cone of $\Fan$,}\end{array}
\\[.7cm]
\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \b{h}_{\ray[s]} > 0
& \begin{array}{l} \text{for any adjacent maximal cones} \\ \text{$\R_{\ge0}\rays$ and~$\R_{\ge0}\rays'$ of~$\Fan[G]$ belonging}\\\text{to \textbf{distinct} maximal cones of $\Fan$}\end{array}
\end{array}
\right\}.
\end{align*}
\end{proposition}

@@

\begin{definition}
\label{def:uerp}
We say that two exchangeable rays~$\ray, \ray'$ of~$\Fan$ admit a \defn{unique exchange relation} when the linear dependence
\[
\sum_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \ray[s] = 0.
\]
does not depend on the pair~$\{\rays, \rays'\}$ of adjacent maximal cones of~$\Fan$ with $\rays \ssm \{\ray\} = \rays' \ssm \{\ray'\}$, but only on the pair of rays~$\{\ray, \ray'\}$.
This implies in particular that the rays~$\ray[s]$ for which~$\coefficient[{\ray[s]}][\rays][\rays'] \ne 0$ belong to~$\rays \cup \rays'$ for any pair~$\{\rays, \rays'\}$ of adjacent maximal cones of~$\Fan$ with $\rays \ssm \{\ray\} = \rays' \ssm \{\ray'\}$.

@@

\begin{definition}
In a fan~$\Fan$ with the unique exchange relation property, an \defn{extremal exchangeable pair} is a pair of exchangeable rays~$\{\ray,\ray'\}$ such that the corresponding inequality~${\sum_{\ray[s]} \coefficient \, \b{h}_{\ray[s]} \ge  0}$ defines a facet of the closed type cone~$\ctypeCone(\Fan)$.
\end{definition}

@@

We still consider an essential complete simplicial fan~$\Fan$ in~$\R^n$, the $N \times n$-matrix~$\b{G}$ whose rows are the rays of~$\Fan$, and the $(N-n) \times N$-matrix~$\b{K}$ which spans the left kernel of~$\b{G}$ (\ie $\b{K}\b{G} = 0$).

@@

\begin{corollary}
\label{coro:simplicialTypeCone}
Assume that the type cone~$\typeCone(\Fan)$ is simplicial and let~$\b{K}$ be the $(N-n) \times N$-matrix whose rows are the inner normal vectors of the facets of~$\typeCone(\Fan)$. Then the polytope
\[
R_\b{\ell} \eqdef \set{\b{z} \in \R^N}{K\b{z} = \b{\ell} \text{ and } \b{z} \ge 0}
\]
is a realization of the fan~$\Fan$ for any positive vector~$\b{\ell} \in \R_{>0}^{N-n}$.
Moreover, the polytopes~$R_\b{\ell}$ for~$\b{\ell} \in \R_{>0}^{N-n}$ describe all polytopal realizations of~$\Fan$.
\end{corollary}

\begin{proof}
Let~$\b{\ell} \in \R_{>0}^{N-n}$.
Since~$\b{K}$ has full rank there exists~$\b{h} \in \R^N$ such that~$\b{K}\b{h} = \b{\ell}$.
Since~$\b{K}\b{h} \ge 0$ and the rows of~$\b{K}$ are precisely all inner normal vectors of the facets of the type cone~$\typeCone(\Fan)$, we obtain that~$\b{h}$ belongs to~$\typeCone(\Fan)$.
Since $R_\b{\ell} = Q_\b{h} \sim P_\b{h}$ by \cref{prop:alternativePolytopalRealization}, we conclude that~$R_\b{\ell}$ is a polytopal realization of~$\Fan$.
Since~$\typeCone(\Fan)$ is simplicial, we have~$\b{K}\typeCone(\Fan) = \R_{>0}^{N-n}$, so that we obtain all polytopal realizations of~$\Fan$ in this way.
\end{proof}

@@

We say that~$P$ is a \defn{Minkowski summand} of~$R$ if there is a polytope~$Q$ such that $P+Q = R$, and a \defn{weak Minkowski summand} if there is a $\lambda \ge 0$ and a polytope~$Q$ such that ${P + Q = \lambda R}$. 
\cref{lem:MinkowskiSum} ensures that convex combinations in the type cone correspond to Minkowski combinations of polytopes. Actually, the set of weak Minkowski summands of a polytope $P$ with normal fan~$\Fan$ is isomorphic to the closed type cone~$\ctypeCone(\Fan)$ modulo linearity (the cone of weak Minkowski summands was studied by W.~Meyer in~\cite{Meyer}, to see that it is equivalent to the type cone and other formulations, see for example the appendix of~\cite{PostnikovReinerWilliams}).

@@

If $P = Q + R$, then we say that~$R$ is the \defn{Minkowski difference} of~$P$ and~$Q$, denoted~${R = P - Q}$. Note that this is only defined when~$Q$ is a Minkowski summand of~$P$.
For instance, we have ${P_\b{h} - P_{\b{h}'} = P_{\b{h} - \b{h'}}}$ when $P_{\b{h}'}$ is a Minkowski summand of $P_\b{h}$.
There are two natural ways this construction generalizes to arbitrary pairs of polyhedra.
On the one hand, one can consider that ${P - Q = \set{x \in \R^n}{x + Q \subseteq P}}$.
On the other hand, considering arbitrary differences of support functions gives rise to \defn{virtual polyhedra}, which is the Grothendieck group of the semigroup of polytopes with Minkowski sums~\cite{PukhlikovKhovanskii}.

@@

\begin{corollary}
If the type cone~$\typeCone(\Fan)$ of $\Fan$ is simplicial, then every $P_\b{h}$ with $\b{h}\in \typeCone(\Fan)$ (resp.~with $\b{h}\in \ctypeCone(\Fan)$) has a unique representation, up to translation, as a Minkowski sum of positive (resp.~non-negative) dilates of the $N-n$ indecomposable polytopes $P_{\b{h}^{(1)}}, \dots, P_{\b{h}^{(N-n)}}$ arising from the rays~$\b{h}^{(1)}, \dots, \b{h}^{(N-n)}$ of $\ctypeCone(\Fan)$.
\end{corollary}

@@

\begin{example}
\label{exm:typeConeMinkowskiSum}
Consider the $2$-dimensional fan~$\Fan$ of \cref{fig:typeCone}\,(left) whose type cone has been described in \cref{exm:typeCone} and represented in \cref{fig:typeCone}\,(right).
As illustrated in \cref{fig:typeConeCoarsenings}, the rays of the closed type cone~$\ctypeCone(\Fan)$ are directed by the height vectors~$(1,2,1,2,1)$ (corresponding to a triangle), $(0,1,1,0,1)$ (corresponding to a vertical segment) and~$(1,0,1,1,0)$ (corresponding to a horizontal segment).
Therefore, any polytope in~$\ctypeCone(\Fan)$ is (up to translation) a positive Minkowski combination of a triangle and two edges.
\cref{fig:typeConeMinkowskiSum} illustrates this property for the polytope of \cref{fig:typeCone}\,(middle) corresponding to the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$.
Note that~$\Fan$ is the normal fan of J.-L.~Loday's $2$-dimensional associahedron, whose realization as a Minkowski sum of faces of a simplex was given by A.~Postnikov in~\cite{Postnikov}.

@@

\begin{theorem}
\label{thm:associahedronLoday}
The associahedron~$\Asso[n]$ can be described equivalently as:
\begin{itemize}
\item the convex hull of the points~$L(T) \in \R^{n+1}$ for all rooted binary trees~$T$ with~$n+1$ internal nodes, where the $i$th coordinate of~$L(T)$ is the product of the number of leaves in the left subtree by the number of leaves in the right subtree of the $i$th node of~$T$ in inorder~\cite{Loday},
\item the intersection of the hyperplane~$\bigset{\b{x} \in \R^{n+1}}{\sum_{\ell \in [n+1]} x_\ell = \binom{n+2}{2}}$ with the halfspaces $\bigset{\b{x} \in \R^{n+1}}{\sum_{a < \ell < b} x_\ell \ge \binom{b-a+2}{2}}$ for all internal diagonals~$(a,b)$ of \mbox{the $(n+3)$-gon~\cite{ShniderSternberg}},
\item the Minkowski sum of the faces~$\simplex_{[i,j]}$ of the standard $n$-dimensional simplex~$\simplex_{[n+1]}$ corresponding to all proper intervals~$[i,j]$ of~$[n+1]$,~\cite{Postnikov}.
\end{itemize}
\end{theorem}

@@

\begin{corollary}
\label{coro:typeConeAsso}
Let~$n \in \N$ and~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$. Then the type cone of the normal fan~$\gvectorFan[n]$ of~$\Asso[n]$ is given by
\[
\typeCone \big( \gvectorFan[n] \big) = \set{\b{h} \in \R^{X(n)}}{\begin{array}{@{}l@{}} \b{h}_{(0,n+2)} = 0, \qquad\text{and}\qquad \b{h}_{(a,a+1)} = 0 \text{ for all } 0 \le a \le n+1 \\ \b{h}_{(a,b)} + \b{h}_{(a',b')} > \b{h}_{(a,b')} + \b{h}_{(a',b)} \text{ for all } 0 \le a < a' < b < b' \le n+2 \end{array}} \!.
\]
\end{corollary}

@@

\medskip
\noindent
In all our tables, we just record the coefficients of the inequalities, which are then easily reconstructed.
For instance, the inequality~$\red \circled{A}$ above is given by~$\b{h}_{\includegraphics[scale=.3]{diagonalA1}} + \b{h}_{\includegraphics[scale=.3]{diagonalA3}} > \b{h}_{\includegraphics[scale=.3]{diagonalA2}}$.

The type cone of~$\gvectorFan[2]$ lives in~$\R^9$ and has a linearity space of dimension~$3$.
It has $6$ facet-defining inequalities (given below), which correspond to the flips described in \cref{prop:extremalExchangeablePairsAsso} and illustrated in \cref{fig:labelFacetDefiningInequalititiesAsso}\,(right).

@@

@@

\begin{corollary}
\label{coro:simplicialTypeConeAsso}
The type cone~$\typeCone \big( \gvectorFan[n] \big)$ is simplicial.
\end{corollary}

@@

\para{Mutations in the cluster fan}
%
We now discuss the linear dependences between $\b{g}$-vectors of adjacent seeds.
The following lemma was stated in~\cite[Lem.~19]{HohlwegPilaudStella}.
